---
title: "All Evenness by Species"
author: "Christopher Bolte"
date: "May 17, 2018"
output: html_document
---

Load the following packages:

```{r}
library(dplyr)
library(ggplot2)
library(grid)
library(gridExtra)
library(vegan)
library(lubridate)
library(readr)
library(gglorenz)
library(ineq)
```

```{r}
#library(readr)
#new_det <- select(det, DetectDate, Detection_GPS, TagID, Location, Species, Release_Location, Release_Date, Release_GPS, Release_rkm = Release_RKm, Detect_rkm = Detection_RKm, Length_Type, Tag_Type)

```

The order of the following functions goes as follows, and can be seperated into the following steps.

1. Get the motilities of all species of fish above a specified length.
  Function: get_species_motility(detection_df, species_id, minimum_fish_length)
    #Uses the functions: determine_years(), get_all_motilities_years() (uses get_year_motility(), add_rank())
  
  This generates a motility data.frame that will be used in other functions.
  
2. Use smush_motility to reorder ranking system to account for all individuals across ALL years instead of for each individual year.
  Function: smush_motility(df_motility)
    #uses add_rank() function.
    
3. Use all_years_evenness() function to generate plots for each species, saving each plot to a simple variable (i.e letters a-e),
then combine these variables into a list, the use the do.call() function with grid.arrange to plot all of them on the same plot.
  #all_years_evenness uses lm_eqn() function



**Load the Following Functions Before Running Example Code**



```{r}
get_year_motility <- function(df, low_rkm=0, high_rkm=550, species_name, str_startdate='2000-01-01 00:00:00', str_enddate='2100-01-01 00:00:00', study_year_month) {
  det <- df[df$Detect_rkm > low_rkm & df$Detect_rkm < high_rkm, ] #Takes detections only within specified rkms
  det <- df[df$DetectDate > as.character(str_startdate) & df$DetectDate < as.character(str_enddate), ] #takes detections only within certain dates #Maybe change as.character back?
  #GLOBALdet <<- det
  df_allmoves <- NULL
  print(nrow(det))
  if (nrow(det) > 0) {
    det$Detect_rkm <- as.numeric(det$Detect_rkm)
    det$Release_rkm <- as.numeric(det$Release_rkm)
    det <- det[is.na(det$Detect_rkm) == FALSE, ]
    #det <- det[is.na(det$Release_rkm) == FALSE, ]
    number_detections <- nrow(det) #number of detections in a single year
    all_tags <- c(unique(det$TagID))
    number_fish <- length(all_tags) #number of individuals detected in a single year. Make sure that these are fish above the specifed length!!!
    
    count <- 1
    total_count <- length(all_tags)
    print(species_name)
    print(all_tags)
    for (tag in all_tags) {
      #Get fish that have been detected for over a specified number of years.
      print(paste(count, '/', total_count))
      ind_tag <- det[det$TagID == tag, ]
      #print(ind_tag)
      #print(nrow(ind_tag))
      if (nrow(ind_tag) > 0) {
        ind_tag <- ind_tag[order(ind_tag$DetectDate), ]
        first_det <- min(ind_tag$DetectDate)
        last_det <- max(ind_tag$DetectDate)
        length <- ind_tag$Length[1]
        all_rows <- c(1:nrow(ind_tag))
        release_rkm <- ind_tag$Release_rkm[1]
        total_rkm <- 0
        for (row in all_rows) {
          first_rkm <- ind_tag$Detect_rkm[row]
          second_rkm <- ind_tag$Detect_rkm[row + 1]
          if (row == nrow(ind_tag)) { #may need to add some subtleties to make sure that if the last detection is different from the second to last that this gets accounted for.
            species <- species_name
            last_location <- ind_tag$Location[row]
            df_allmoves <- rbind(df_allmoves, data.frame(study_year_month, tag, total_rkm, species, length, first_det, last_det, number_fish, number_detections, last_location))
            break
          }
          if (first_rkm != second_rkm) {
            diff_rkm <- abs(first_rkm - second_rkm)
            total_rkm <- total_rkm + diff_rkm
          }
        }
        count <- count + 1
      }
    }
  }
  #Add a statement that will add bins to the analysis, will need to change how it makes the ggplots in other functions
  return(df_allmoves)
}

```


```{r}
add_rank <- function(df) {
  det <- df
  det <- det[order(det$total_rkm, decreasing = TRUE), ]
  det$index <- c(1:nrow(det))
  return(det)
}

```

```{r}
determine_years <- function(df) {
  test_df <- df
  test_df$DetectDate <- as.Date(test_df$DetectDate, format = '%y')
  test_df$DetectDate <- year(as.Date(test_df$DetectDate, '%y-%m-%d'))
  years <- c(unique(test_df$DetectDate))
  df_all_years <- NULL
  for (y in years) {
    ind_yearr <- df_years[df_years$year == y, ]
    if (nrow(ind_yearr) > 0) {
     df_all_years <- rbind(df_all_years, ind_yearr)
    }
  }
  df_all_years <- df_all_years[order(df_all_years$year, decreasing = FALSE), ]
  return(df_all_years)
}
```


```{r}
get_all_motilities_years <- function(df, species_name, year_list) {
  #Year list should have 3 entries, 'year', 'start', 'end'. This should be made before the loop
  df_all_motilities <- NULL
  all_years <- 1:nrow(year_list)
  for (y in all_years) { #y should be a number index
    start_datetime <- as.character(year_list$start_date[y])
    end_datetime <- as.character(year_list$end_date[y])
    year_month <- as.character(year_list$year[y])
    print(start_datetime)
    print(end_datetime)
    print(year_month)
    motility <-  df %>% 
      get_year_motility(species_name = species_name, str_startdate = start_datetime, str_enddate = end_datetime, study_year_month=year_month) %>%
      add_rank
    df_all_motilities <- rbind(df_all_motilities, motility)
  }
  return(df_all_motilities)
}

```



```{r}

lm_eqn <- function(df){
  new_df <- df
  #new_df <- subset(new_df, proportion > 0.0000)
  m <- lm(proportion ~ rank, new_df);
  eq <- substitute(italic(proportion) == a + b %.% italic(rank)*","~~italic(r)^2~"="~r2, 
                   list(a = format(coef(m)[1], digits = 2), 
                        b = format(coef(m)[2], digits = 2), 
                        r2 = format(summary(m)$r.squared, digits = 3)))
  as.character(as.expression(eq));                 
}
```



```{r}
get_species_motility <- function(df, species_id, int_length) { ###Add this to Markdown Documents !!!
  ind_species <- subset(df, Species == species_id)
  ind_species <- subset(ind_species, is.null(ind_species$Length) == FALSE)
  ind_species <- subset(ind_species, Length != 'NULL')
  ind_species$Length <- as.numeric(ind_species$Length)
  ind_species <- subset(ind_species, Length > int_length)
  df_year_info <- determine_years(ind_species)
  all_motilities <- ind_species %>% get_all_motilities_years(species_id, df_year_info) %>%  purge_ggfish() %>%  add_rank_byyear()
  return(all_motilities)
}

```


```{r}
add_rank_byyear <- function(df) { ### Add to Markdown Documents
  det <- df
  new_df <- NULL
  all_years <- unique(det$study_year_month)
  for (y in all_years) {
    print(y)
    ind_year <- subset(det, study_year_month == y)
    ind_year <- ind_year[order(ind_year$total_rkm, decreasing = TRUE), ]
    ind_year$index <- c(1:nrow(ind_year))
    new_df <- rbind(new_df, ind_year)
  }
  return(new_df)
}
```


```{r}
all_years_evenness <- function(df, title) { #df should be a motility data.frame that has gone through the smush_motility function.
  motility <- df
  
  #remove values that will log transform to infinity
  #motility <- subset(motility, proportion > 0.0000)
  
  motility$proportion <- signif(motility$proportion, digits = 5)
  df_shannon <- motility %>% group_by(proportion) %>% tally()
  specnumber <- length(unique(df_shannon$proportion))
  shannon_index <- diversity(df_shannon$n)
  pielou_index <- signif(shannon_index/(specnumber), digits = 3)
  print(pielou_index)
  
  ggplot(data = motility, aes(x=rank, y= proportion)) + 
    geom_point() + 
    geom_line() +
    theme(legend.position = 'none') +
    xlab('Rank') +
    ylab('Proportion') +
    ggtitle(title) +
    #ggtitle(paste(study_year, 'Rank vs Motility of', species_name, 'in Sacramento/San Joaquin Watershed')) +
    theme(plot.title = element_text(size = 10, face = "bold")) +
    theme(axis.title = element_text(size=8)) +
    coord_cartesian(ylim = c(0, 1)) + 
    geom_smooth(method='lm', col='red', size=.25, se = FALSE) +
    annotate('text', x=nrow(motility)/4, y=.5, label = paste('Index of Evenness:', pielou_index, sep = ' '), size = 2.5) + 
    geom_text(x = nrow(motility)/3, y = .75, label = lm_eqn(motility), parse = TRUE, size = 2.5)
}

```

```{r}
smush_motility <- function(df) {
  mot <- df
  mot$index <- NULL
  new_mot <- add_rank(mot)
  return(new_mot)
}
```

```{r}
get_proportionality_total <- function(df) {
  df_mot <- df
  df_mot$proportion <- NA
  df_mot$rank <- NA
  all_rows <- c(1:nrow(df_mot))
  all_rkms <- sum(df_mot$total_rkm)
  for (row in all_rows) {
    x <- df_mot$total_rkm[row]
    df_mot$proportion[row] <- x/all_rkms
  }
  df_mot <- df_mot[order(df_mot$proportion, decreasing = TRUE), ]
  df_mot$rank <- c(1:nrow(df_mot))
  return(df_mot)
}
```

```{r}
purge_ggfish <- function(df, min_rkm = 2) {
  new_mot <- df[!(df$total_rkm <= min_rkm & substr(df$last_location, 1, 2) == 'GG'), ]
  return(new_mot)
}
```


```{r}
all_years_evenness_gini <- function(df, title) { #df should be a motility data.frame that has gone through the smush_motility function.
  df_prop <- df
  
  #remove values that will log transform to infinity
  #df_prop <- subset(df_prop, proportion > 0.0000)
  
  gini_index <- signif(ineq(df_prop$proportion, type = 'Gini'), digits = 3)
  
  ggplot(data = df_prop, aes(x=rank, y= proportion)) + 
    geom_point() + 
    geom_line() +
    theme(legend.position = 'none') +
    xlab('Rank') +
    ylab('Proportion') +
    ggtitle(title) +
    #ggtitle(paste(study_year, 'Rank vs Proportion', species_name, 'in Sacramento/San Joaquin Watershed')) +
    theme(plot.title = element_text(size = 10, face = "bold")) +
    theme(axis.title = element_text(size=8)) +
    coord_cartesian(ylim = c(0, .1)) + 
    geom_smooth(method='lm', col='red', size=.25, se = FALSE) +
    annotate('text', x=nrow(df_prop)/4, y=.025, label = paste('Gini Coefficient:', gini_index, sep = ' '), size = 2.5) + 
    geom_text(x = nrow(df_prop)/3, y = .75, label = lm_eqn(df_prop), parse = TRUE, size = 2.5)
}
```

```{r}
get_lorenz_interspecies <- function(df_proportion, species_name) { 
  df_prop <- df_proportion
  gini_index <- signif(ineq(df_prop$proportion), digits = 3)
  print(gini_index)
  p <- ggplot(data = df_prop, aes(proportion)) +
    ggtitle(species_name) +
    stat_lorenz(desc = FALSE) +
    coord_fixed() +
    geom_abline(linetype = "dashed") +
    theme(axis.title = element_text(size = 6), plot.title = element_text(size = 8)) +
    annotate('text', x = .25, y = .75, label = paste('Gini Index', gini_index), size = 2)
  return(p)
}
```

```{r}
#species_name should be any of 'Green Sturgeon', 'White Sturgeon', 'Sacramento Pikeminnow', 'Steelhead', 'Striped Bass'
all_years_evenness_species <- function(df, species_name) { #df should be a motility data.frame that has gone through the smush_motility function.
  
  #Will need to adjust position of Pielous index based on SPECIES
  
  df_prop <- df
  species <- species_name
  species_color <- NA
  if (species_name == 'Green Sturgeon') { #Could add in a position for where Pielou's goes?
    species_color <- 'mediumaquamarine'
    p_location <- 110
    ybreak <- c(0, .025, .05)
    ylimit <- c(0, .05)
  }
  if (species_name == 'White Sturgeon') {
    species_color <- 'darkgoldenrod1'
    p_location <- 18
    ybreak <- c(0, .10, .20)
    ylimit <- c(0, .20)
  }
  if (species_name == 'Sacramento Pikeminnow') {
    species_color <- 'black'
    p_location <- 50
    ybreak <- c(0, .25, .5)
    ylimit <- c(0, .5)
  }
  if (species_name == 'Steelhead') {
    species_color <- 'deepskyblue2'
    p_location <- 50
    ybreak <- c(0, .035, .075)
    ylimit <- c(0, .075)
  }
  if (species_name == 'Striped Bass') {
    species_color <- 'orangered2'
    p_location <- 9
    ybreak <- c(0, .03, .06)
    ylimit <- c(0, .06)
  }
  
  select_df <- select(df_prop, study_year_month, total_rkm)
  specnumber <- length(unique(select_df$t))
  shannon_index <- diversity(select_df$total_rkm)
  pielou_index <- signif(shannon_index/(log(specnumber)), digits = 2)
  
  ggplot(data = df_prop, aes(x=rank, y= proportion)) + 
    geom_line(size = 2, color = species_color) +
    theme(legend.position = 'none') +
    #ggtitle(species) +
    #ggtitle(paste(study_year, 'Rank vs df_prop of', species_name, 'in Sacramento/San Joaquin Watershed')) +
    #theme(plot.title = element_text(size = 4, face = "bold")) +
    scale_y_continuous(breaks = ybreak) +
    coord_cartesian(ylim = ylimit, xlim <- c(1,nrow(df_prop))) + 
    annotate('text', x=nrow(df_prop)/4, y=ylimit[2]*.9, label = paste("Pielou's Index:", pielou_index, sep = ' '), size = 3.5) + #change text on plot
    theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank())
}

```

```{r}
get_lorenz_interspecies <- function(df_proportion, species_name) { 
  df_prop <- df_proportion
  species <- species_name
  species_color <- NA
  if (species_name == 'Green Sturgeon') { #Could add in a position for where Pielou's goes?
    species_color <- 'mediumaquamarine'
  }
  if (species_name == 'White Sturgeon') {
    species_color <- 'darkgoldenrod1'
  }
  if (species_name == 'Striped Bass') {
    species_color <- 'orangered2'
  }
  if (species_name == 'Sacramento Pikeminnow') {
    species_color <- 'black'
  }
  if (species_name == 'Steelhead') {
    species_color <- 'deepskyblue2'
  }
  gini_index <- signif(ineq(df_prop$proportion), digits = 2)
  print(gini_index)
  p <- ggplot(data = df_prop, aes(proportion)) +
    #ggtitle(species_name) +
    stat_lorenz(desc = FALSE, geom = 'line', color=species_color, size = 2) +
    coord_fixed(1/4) +
    geom_abline(linetype = "dashed") +
    scale_y_continuous(breaks = c(0.0, 0.5, 1.0)) +
    annotate('text', x = .22, y = .9, label = paste("Gini's Index", gini_index), size = 3.5) + #change text on plot
    theme(plot.title = element_text(size = 6)) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(axis.title.x=element_blank(), axis.title.y = element_blank()) 
  
    #theme(plot.margin = unit(c(0, 3, .2, .2), 'cm'))   # top, 
  return(p)
}

```


Runthrough Code:

```{r}
#Get Interspecies Plots along with Gini Index and Loren Graphs


#Import Each Species Detections (Filtered)
gs_det <- read_csv("~/Desktop/TRUE FINAL EVENNESS PROJECT/CSVs/Length Adjusted CSVs/Final_Green_Sturgeon.csv", 
                   col_types = cols(DetectDate = col_datetime(format = "%Y-%m-%d %H:%M:%S")))

ws_det <- read_csv("~/Desktop/TRUE FINAL EVENNESS PROJECT/CSVs/Length Adjusted CSVs/Final_White_Sturgeon.csv", 
                   col_types = cols(DetectDate = col_datetime(format = "%Y-%m-%d %H:%M:%S")))

pike_det <- read_csv("~/Desktop/TRUE FINAL EVENNESS PROJECT/CSVs/Length Adjusted CSVs/Final_Pikeminnow.csv", 
                   col_types = cols(DetectDate = col_datetime(format = "%Y-%m-%d %H:%M:%S")))

sth_det <- read_csv("~/Desktop/TRUE FINAL EVENNESS PROJECT/CSVs/Length Adjusted CSVs/Final_Steelhead.csv", 
                   col_types = cols(DetectDate = col_datetime(format = "%Y-%m-%d %H:%M:%S")))

sb_det <- read_csv("~/Desktop/TRUE FINAL EVENNESS PROJECT/CSVs/Length Adjusted CSVs/Final_Striped_Bass.csv", 
                   col_types = cols(DetectDate = col_datetime(format = "%Y-%m-%d %H:%M:%S")))

#check to make sure import went well
head(gs_det)
head(ws_det)
head(pike_det)
head(sth_det)
head(sb_det)

df_years <- read_csv("~/Desktop/TRUE FINAL EVENNESS PROJECT/CSVs/water_year_data.csv", 
                     col_types = cols(end_date = col_character(), 
                                      start_date = col_character()))

all_det <- rbind(gs_det, pike_det, sb_det, sth_det, ws_det)

det <- select(all_det, DetectDate,TagID, Location, Species, Release_Date, Release_rkm, Detect_rkm, Length, Length_Type, Tag_Type)
#det <- subset(det, Species == 'White_Sturgeon' | Species == 'Striped_bass' | Species == 'Sacramento_pikeminnow' | Species == 'Steelhead')


#Get Motilities
gsturg_motility <- get_species_motility(det, 'Green_Sturgeon', 1800)
wsturg_motility <- get_species_motility(det, 'White_Sturgeon', 1600)
striped_motility <- get_species_motility(det, 'Striped_bass', 440)
pike_motility <- get_species_motility(det, 'Sacramento_pikeminnow', 310)
steelhead_motility <- get_species_motility(det, 'Steelhead',350)

#Condense Detections and Convert To Proportionality
all_green <- smush_motility(gsturg_motility) %>% get_proportionality_total()
all_white <- smush_motility(wsturg_motility) %>% get_proportionality_total()
all_striped <- smush_motility(striped_motility) %>% get_proportionality_total()
all_pike <- smush_motility(pike_motility) %>% get_proportionality_total()
all_steelhead <- smush_motility(steelhead_motility) %>% get_proportionality_total()

#Assemble Rank/Pielous Plots of All Species

a <- all_years_evenness_species(all_green, 'Green Sturgeon')
b <- all_years_evenness_species(all_white, 'White Sturgeon')
c <- all_years_evenness_species(all_pike, 'Sacramento Pikeminnow')
d <- all_years_evenness_species(all_steelhead, 'Steelhead')
e <- all_years_evenness_species(all_striped, 'Striped Bass')

#Create Lorenz/Gini Plots for All Species
f <- get_lorenz_interspecies(all_green, 'Green Sturgeon')
g <- get_lorenz_interspecies(all_white, 'White Sturgeon')
h <- get_lorenz_interspecies(all_pike, 'Sacramento Pikeminnow')
i <- get_lorenz_interspecies(all_steelhead, 'Steelhead')
j <- get_lorenz_interspecies(all_striped, 'Striped Bass')

#Combine Pairs based on species
aa <- grid.arrange(a, f, ncol=2, top = textGrob('Green Sturgeon', gp=gpar(fontface='bold',fontsize=10)))
bb <- grid.arrange(b, g, ncol=2, top = textGrob('White Sturgeon', gp=gpar(fontface='bold',fontsize=10)))
cc <- grid.arrange(c, h, ncol=2, top = textGrob('Sacramento Pikeminnow', gp=gpar(fontface='bold',fontsize=10)))
dd <- grid.arrange(d, i, ncol=2, top = textGrob('Steelhead', gp=gpar(fontface='bold',fontsize=10)))
ee <- grid.arrange(e, j, ncol=2, top = textGrob('Striped Bass', gp=gpar(fontface='bold',fontsize=10)))

#Make Final Plots
grid.arrange(aa, cc, dd, ee,bb, ncol=1, nrow=5, left = 'Total Movement', 
             bottom='Rank           Cumulative % of Movement Observations') #Final Product

## X Axis labels will need to be adjusted manually.

```








